### 2016-05-12 Draft.tex
A few revisions to introduction for readability and flow.  A few comments to self added elsewhere.

\adam{WRITE A SPELLBINDING OPENING PARAGRAPH}\jarad{no need to aim so high}

@@

Removal sampling is a species abundance surveying methodology in which observers capture and physically remove animals at a survey site over a series of trapping sessions.  
Assuming that the probability of capturing a not-yet-trapped individual is the same for all trapping sessions, it is possible to estimate the proportion of individuals still not captured from the pattern of captures over time \citep{Moran1951, Zippin1958, Seber1982}.  
\citet{Farnsworth2002} adapted removal sampling to avian point-count surveys, suggesting that observers record just the first time-to-detection for each bird and mentally remove it thereafter.  
Analysis proceeds by subdividing the observation period into intervals of equal duration and assuming that the rate of detection is the same at all times (i.e., following a Poisson process); then, each interval can be treated as a trapping event, and the interval-specific probability of detection can be estimated as with conventional removal sampling.

However, the assumption of a constant detection rate throughout the observation period is not always justified \citep{Alldredge2007}.  
Animal behaviors such as bout singing or intermittent singing in birds and frogs or diving in whales can lead to time-varying rates of detection \citep{Scott2005, Diefenbach2007, Reidy2011}.  
The act of observation may affect animal behavior.  
Depending upon the study species, the presence of an observer may temporarily either stimulate or suppress detectable cues \citep{McSheaRappole1997, Rosenstock2002, Alldredge2007}.  
Survey methodology itself can affect detection rates.  
Observers may also become distracted as the survey progresses \citep{Johnson2008}.  
And, depending upon protocol, observers may become aware of animals during their settling-in period, resulting in elevated counts early during in the survey period \citep{LeeMarsden2008}.  
In the presence of detection heterogeneity, easily detected animals will be removed first, leaving only harder to detect animals, thus resulting in a marginal detection rate that varies over time \citep{Farnsworth2005}.

@@

@@

The utility of applying a mixture TTDD for constant-detection models has been identified several times \citep{Pledger2000, Farnsworth2002, Alldredge2007, Reidy2011}.  
The mixture allows an analysis to account for detection heterogeneity across easy- and hard-to-detect subgroups of a population.  
Even so, some recent studies have assumed the absence of a mixture \citep{Solymos2013, Amundson2014}.  
Our study indicates that mixtures are useful even when a non-constant TTDD is used for hard-to-detect individuals.  
Models that include a mixture generally perform well even when no heterogeneity is present, but models that omit a mixture are biased and overly certain when detection heterogeneity exists.  
Indeed, for peaked datasets, mixture models can outperform non-mixture models even in the absence of heterogeneity.

@@

The time-to-detection approach in our model is a departure from the probability-of-detection approach that has been used in distance-removal modeling \citep{Farnsworth2005, Diefenbach2007, Solymos2013, Amundson2014}\adam{Alldredge2007?}.  
In a distance-sampling context, the probability of detection for any individual is framed as the joint probability of two independent detection events: (i) availability for detection, which occurs during an observation period with probability $P_a$, and (ii) detection of the available individual by an observer, $P_b$, also known as perceptibility \citep{Williams2002, Kery2008, Nichols2009}\adam{I haven't read Williams2002 -- cited in Royle2004.  
I'm  beginning to think this may have too-numerous antecedents to cite... in Reidy, the earliest reference is Marsh and Sinclair 1989.  
See also McCallum 2005.}.  
In avian point-counts, because the overwhelming majority of detections are auditory, $P_a$ is understood to be the probability that a bird vocalizes during the observation period.  
Meanwhile, perceptibility is treated as a function of distance $d$: $P_b = g(d)$, with nearby available individuals being detected with higher probability than distant individuals.  
The probability of detection is thus formulated as availability times perceptibility: $p^{(det)} = P_aP_b$.  
Multiple studies have demonstrated that failure to incorporate distance leads to systemic bias in estimates of abundance \citep{EffordDawson2009, Solymos2013}.

>>>>

Abundance estimates from animal point-count surveys require accurate estimates of detection probabilities.  
Removal sampling provide adapted to point counts provides one established methodology for estimating detection probabilities.
% cost-effective in that it requires neither multiple observers nor multiple site visits.
Removal sampling is a species abundance surveying methodology in which observers capture and physically remove animals at a survey site over a series of trapping sessions.
Assuming that individual capture rates remain constant over all trapping sessions, it is possible to estimate the proportion of individuals still not captured from the pattern of captures over time .  
\citet{Farnsworth2002} adapted removal sampling to point-count surveys, suggesting that if: (i) observers record only the first time-to-detection for each animal, and (ii) times to first detection are interval-censored, then data will be similar in for to traditional removal sampled data.
In the analysis of such data, the observation period is subdivided into unit intervals of equal duration, each of which is equivalent to a trapping session.

@@

However, the assumption of a constant detection rate throughout the observation period is not always justified \citep{Alldredge2007}. \adam{Others, like Mantyniemi.}  
Animal behaviors such as intermittent singing in birds and frogs or diving in whales can lead to time-varying rates of detection \citep{Scott2005, Diefenbach2007, Reidy2011, Borcherspre-print}.  
The act of observation may affect animal behavior.  
The presence of an observer may temporarily either stimulate or suppress detectable cues \citep{McSheaRappole1997, Rosenstock2002, Alldredge2007}.  
Survey methodology itself can affect recording of detection times.  
Observers may become aware of animals during their settling-in period, resulting in elevated counts early during in the survey period \citep{LeeMarsden2008}.  
Observers may also become distracted as the survey progresses \citep{Johnson2008}.  
In the presence of detection heterogeneity, easily detected animals will be removed first, leaving only harder to detect animals, thus resulting in a marginal detection rate that varies over time \citep{Farnsworth2005}.

@@

%Application of this methodology generally assumes that organisms at a survey site are detected at a constant rate \citep{Farnsworth2005, Etterson2009, Solymos2013, Amundson2014, Reidy2016}; however, this assumption is often not justified in the field.  
%Failure to account for non-constant detection leads to biased estimates of detection and therefore of abundance.

@@

The utility of applying a mixture TTDD for constant-detection models has been identified several times \citep{Pledger2000, Farnsworth2002, Alldredge2007, Reidy2011}.  
The mixture allows an analysis to account for detection heterogeneity across easy- and hard-to-detect subgroups of a population.  
Even so, some recent studies have assumed the absence of a mixture \citep{Solymos2013, Amundson2014}.  
Our study indicates that mixtures are useful even when a non-constant TTDD is used for hard-to-detect individuals.  
Models that include a mixture generally perform well even when no heterogeneity is present, but models that omit a mixture are biased and overly certain when detection heterogeneity exists.  
Indeed, for peaked datasets, mixture models can outperform non-mixture models even in the absence of heterogeneity.
% "It is tempting to choose classes of models that yield narrower confidence intervals for N; however, Coull's (1997) extensive simulation studies (summarized by Coull and Agresti (1999)) clearly reveal that this behavior can produce erroneous inferences." -- Dorazio and Royle 2003

@@

The time-to-detection approach in our model is a departure from the probability-of-detection approach that has been used in distance-removal modeling \citep{Farnsworth2005, Diefenbach2007, Solymos2013, Amundson2014}\adam{Alldredge2007?}.  
In a distance-sampling context, the probability of detection for any individual is framed as the joint probability of two independent detection events: (i) availability for detection, which occurs during an observation period with probability $P_a$, and (ii) detection of the available individual by an observer, $P_b$, also known as perceptibility \citep{Williams2002, Kery2008, Nichols2009}\adam{I haven't read Williams2002 -- cited in Royle2004.  
I'm  beginning to think this may have too-numerous antecedents to cite... in Reidy, the earliest reference is Marsh and Sinclair 1989.  
Ditto Amundson.
See also McCallum 2005.}.  
In avian point-counts, because the overwhelming majority of detections are auditory, $P_a$ is understood to be the probability that a bird vocalizes during the observation period.  
Meanwhile, perceptibility is treated as a function of distance $d$: $P_b = g(d)$, with nearby available individuals being detected with higher probability than distant individuals.  
The probability of detection is thus formulated as availability times perceptibility: $p^{(det)} = P_aP_b$.  
Multiple studies have demonstrated that failure to incorporate distance leads to systemic bias in estimates of abundance \citep{EffordDawson2009, Solymos2013}.
